---
title: "happign for foresters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{happign for foresters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.height = 4,
  fig.width = 6
)
```

```{r message=FALSE, warning=FALSE}
library(happign)
library(sf)
library(tmap); tmap_mode("view") # Set map to interactive
library(dplyr)
library(ggplot2);sf_use_s2(FALSE) # Avoid problem with spherical geometry
library(purrr)
library(stars)
```

## First choose a zone of interest

For the example we will work with the Camors forest. First of all we need the commune border that can be obtained from the insee code. Fortunately, happign provides a table containing all insee codes (`data("cog_2022")`). Then, `get_apicarto_commune` is used to donxload shape from Apicarto commune.

```{r}
data("cog_2022")

insee_code <- cog_2022[grepl("^Camors", cog_2022$LIBELLE),1]

borders <- get_apicarto_commune(insee_code)

tm_shape(borders)+
   tm_borders()+
   tm_text("nom_com")

```

## Download cadastral parcel

Cadastral parcels are essential for any forest management. Here how to download it with `get_wfs`.

```{r}

apikey <- "parcellaire"
layer_name <- get_layers_metadata(apikey, "wfs")
name_parcellaire_layer <- layer_name[15,1] # "CADASTRALPARCELS.PARCELLAIRE_EXPRESS:parcelle"

parcellaire <- get_wfs(shape = borders,
                       apikey = apikey,
                       layer_name = name_parcellaire_layer) %>% 
   st_transform(st_crs(borders)) %>% 
   st_intersection(borders)

tm_shape(borders)+
   tm_borders(col = "red", lwd = 2)+
tm_shape(parcellaire)+
   tm_polygons(alpha = 0)

```

__*Rq :*__

 * *IGN WFS can return a maximum of 1000 features. The get_wfs() function overrides this limit by performing several consecutive requests, that why there 6 request above ;*
 * *The shape used get_wfs() is converted to bbox before requesting. In other words, several parcels will be outside our zone, which is why an intersection is made.*
 
## Downloading BD Forêt

The first interesting layer for forester is the "BD Forêt" which is all vegetation type assigned to each area greater than or equal to 0.5 ha (5,000 m²). There is two layer for forest : the old one BD Forêt V1 and the new one BD Forêt V2 that can be accessed with "burger menu" on the top left of interactive map below.

```{r warning=FALSE}
apikey <- "environnement"
layer_name <- get_layers_metadata(apikey, "wfs")
name_BDV1 <- layer_name[1,1]
name_BDV2 <- layer_name[2,1]

BDV1 <- get_wfs(borders, apikey, name_BDV1) %>% 
   st_transform(st_crs(borders)) %>% 
   st_intersection(borders)

BDV2 <- get_wfs(borders, apikey, name_BDV2) %>%
   st_transform(st_crs(borders)) %>%
   st_intersection(borders)

tm_shape(BDV1) +
   tm_polygons(col = "libelle",
               popup.vars = names(BDV1)[1:(ncol(BDV1)-2)],
               legend.show = FALSE)+
tm_shape(BDV2) +
    tm_polygons(col = "tfv",
                alpha = 0.5,
                popup.vars = names(BDV2)[1:(ncol(BDV2)-2)],
                legend.show = FALSE) +
tm_shape(borders) +
   tm_borders(lwd = 2)
```

More calculations can be done as you can see below :

```{r warning=FALSE}
forest_type_BDV2 <- BDV2 %>%
  mutate(area = as.numeric(st_area(geometry))) %>%
  st_drop_geometry() %>%
  group_by(essence) %>%
  summarise(sum_area = sum(area)/10000) %>%
  arrange(desc(sum_area)) %>%
  mutate(essence = as.factor(essence))

ggplot()+
  geom_col(data = forest_type_BDV2,
           aes(x = rev(reorder(essence, sum_area)),
               y = sum_area,
               fill = as.factor(essence)))+
  theme_bw()+
  labs(title = "Surface couverte par essences [ha]",
       y = "Surface [ha]",
       fill = "Essence :")+
  theme(axis.text.x = element_blank())


```

## Detect protected area

One information you really want when you work at forest management is if your zone of interest is inside protected area. The example code below is design to automatically test every layer starting with "PROTECTED" so you can be sure that you have all of them.

Again, you can click on map, point and shape for more informations.

```{r message=FALSE, warning=FALSE}
apikey <- "environnement"
protected_area_names <- get_layers_metadata(apikey, "wfs") %>% 
   filter(grepl("^PROTECTED", Name)) %>% 
   pull(Name)

all_protected_area <- map(.x = protected_area_names,
                          .f = ~ try(get_wfs(borders, apikey, .x), silent = TRUE)) %>% 
   set_names(protected_area_names) %>% 
   discard(~ dim(.)[2] == 1)

# Plot the result
tm_shape(all_protected_area[[1]])+
   tm_polygons(group = "APB", alpha = 0.8, col = "blue")+
tm_shape(all_protected_area[[2]])+
   tm_dots(group = "Point rencontre des secours en forêts", col = "red")+
tm_shape(all_protected_area[[3]])+
   tm_polygons(group = "SIC", alpha = 0.8, col = "purple")+
tm_shape(all_protected_area[[4]])+
   tm_polygons(group = "Znieff 2", alpha = 0.8, col = "pink")+
tm_shape(all_protected_area[[5]])+
   tm_polygons(group = "ZPS", alpha = 0.8, col = "green")+
tm_shape(borders,is.master = TRUE) +
   tm_borders(lwd = 2)
```


## MNS, MNT and MNH...

It's always good to know a more about terrain topologie. IGN offers MNT and MNS for download. As a reminder, the MNT corresponds to the surface of the ground and the MNS to the real surface (in our case, the trees). It is thus easy to find the height of the trees by subtracting the DTM from the MNS.

```{r}
layers_name <- get_layers_metadata("altimetrie", "wms")
MNT_name <- layers_name[3,1] # "ELEVATION.ELEVATIONGRIDCOVERAGE.HIGHRES"
MNS_name <- layers_name[4,1] # "ELEVATION.ELEVATIONGRIDCOVERAGE.HIGHRES.MNS"

MNT <- get_wms_raster(borders, "altimetrie", resolution = 5, unlist(MNT_name), filename = "MNT")
MNS <- get_wms_raster(borders, "altimetrie", resolution = 5, unlist(MNS_name), filename = "MNS")

level_curve <- get_wfs(borders, "altimetrie", "ELEVATION.CONTOUR.LINE:courbe") %>% 
   st_intersection(borders)

MNH <- MNS - MNT
MNH[MNH < 0] <- NA  # Remove negative value 
MNH[MNH > 50] <- NA # Remove height more than 50m

tm_shape(MNH)+
   tm_raster()+
tm_shape(level_curve)+
   tm_lines()+
tm_shape(borders)+
   tm_borders(lwd = 2)

file.remove(list.files()[grep("^MNT",list.files())]) # I don't want to keep MNT on disk because this is vignette
file.remove(list.files()[grep("^MNS",list.files())])# I don't want to keep MNS on disk because this is vignette
```

## NDVI

The code below present the calculation of the NDVI. All informations and palette come from this [website](https://www.usna.edu/Users/oceano/pguth/md_help/html/ndvi_colors.html)
The value range of an NDVI is -1 to 1. It is (Near Infrared - Red) / (Near Infrared + Red) :

 * Water has a low reflectance in red, but almost no NIR (near infrared) reflectance.  So the difference will be small and negative, and the sum will be small, and NDVI large and negative.
 * Plants have a low reflectance in red, and a strong NIR reflectance.  So the difference will be large and positive, and the sum will be just about the same as the difference, so NDVI will be large and positive.

Categories are somewhat arbitrary, and you can find various rules of thumb, such as:

 * Negative values of NDVI (values approaching -1) correspond to water. Values close to zero (-0.1 to 0.1) generally correspond to barren areas of rock, sand, or snow. Low, positive values represent shrub and grassland (approximately 0.2 to 0.4), while high values indicate temperate and tropical rainforests (values approaching 1).
 * Very low values of NDVI (0.1 and below) correspond to water, barren areas of rock, sand, or snow. Moderate values represent shrub and grassland (0.2 to 0.3), while high values indicate temperate and tropical rainforests (0.6 to 0.8).


```{r}
# To show the possibility of 20cm resolution for IRC, let's take only the biggest parcels
biggest_parcels <- parcellaire %>% 
   mutate(area = st_area(.)) %>% 
   slice_max(area)

IRC <- get_wms_raster(shape = biggest_parcels,
                      apikey = "ortho",
                      resolution = 0.2,
                      layer_name = "ORTHOIMAGERY.ORTHOPHOTOS.IRC",
                      filename = "IRC")
IRC <- st_as_stars(IRC)

infrared <- IRC %>% slice(band, 1)
red <- IRC %>% slice(band, 2)

NDVI <- (infrared-red)/(infrared+red)
   
breaks_NDVI <- c(-1,-0.2,-0.1,0,0.025 ,0.05,0.075,0.1,0.125,0.15,0.175,0.2 ,0.25 ,0.3 ,0.35,0.4,0.45,0.5,0.55,0.6,1)
   
palette_NDVI <- c("#BFBFBF","#DBDBDB","#FFFFE0","#FFFACC","#EDE8B5","#DED99C","#CCC782","#BDB86B","#B0C261","#A3CC59","#91BF52","#80B347","#70A340","#619636","#4F8A2E","#407D24","#306E1C","#216112","#0F540A","#004500")

tm_shape(borders)+
   tm_borders(lwd = 2, col = "red")+
tm_shape(NDVI)+
   tm_raster(stretch.palette = F,
             breaks = breaks_NDVI,
             palette = palette_NDVI,
             colorNA = "red")+
tm_shape(biggest_parcels)+
   tm_borders(lwd = 2, col = "blue")

file.remove(list.files()[grep("^IRC",list.files())]) # I don't want to keep IRC on disk because this is vignette
```

## The gloss index

The gloss index represents the average of the image glosses. This index is therefore sensitive to the brightness of the soil, related to its moisture and the presence of salts on the surface. It characterizes especially the albedo (solar radiation that is reflected back to the atmosphere). The gloss index allows us to estimate whether the observed surface feature is light or dark.

```{r}

gloss_index <- sqrt(red^2 + infrared^2)

tm_shape(borders)+
   tm_borders(lwd = 2, col = "red")+
tm_shape(gloss_index)+
   tm_raster()+
tm_shape(biggest_parcels)+
   tm_borders(lwd = 2, col = "blue")

```


## Last but not least... BD Topo

BD topo from IGN covers in a coherent way all the geographical and administrative entities of the national territory. So you can find in it :

- Administrative (boundaries and administrative units);
- Addresses (mailing addresses) ;
- Building (constructions) ;
- Hydrography (water-related features) ;
- Named places (place or locality with a toponym describing a natural space or inhabited place);
- Land use (vegetation, foreshore, hedge);
- Services and activities (utilities, energy storage and transportation, industrial sites);
- Transportation (road, rail or air infrastructure, routes);
- Regulated areas (most of the areas are subject to specific regulations).

For the example below I choose to download all water-related data :

```{r}
cour_eau <- get_wfs(borders, "topographie", "BDTOPO_V3:cours_d_eau") %>% 
   st_intersection(borders)

detail_hydro <- get_wfs(borders, "topographie", "BDTOPO_V3:detail_hydrographique")  %>% 
   st_intersection(borders)

# water detected by satellite
surf_hydro <- get_wfs(borders, "topographie", "BDTOPO_V3:surface_hydrographique")  %>% 
   st_intersection(borders)

tm_shape(cour_eau)+
   tm_lines(col = "blue")+
tm_shape(detail_hydro)+
   tm_dots(col = "red")+
tm_shape(surf_hydro)+
   tm_polygons("steelblue")+
tm_shape(borders)+
   tm_borders(lwd = 2)
```

## What about history ?

The "Etat-major" map is a general map of France made, in its first version, in the 19th century. Here how to get it :

```{r}
etat_major <- get_wms_raster(shape = borders,
                            apikey = "cartes",
                            resolution = 1.5,
                            layer_name = "GEOGRAPHICALGRIDSYSTEMS.ETATMAJOR40",
                            filename = "etat_major")

etat_major <- st_as_stars(etat_major, ignore_file = TRUE)

tm_shape(etat_major)+
   tm_rgb()+
tm_shape(borders)+
   tm_borders(lwd = 3, col = "red")

file.remove(list.files()[grep("^etat_major",list.files())])

```
