---
title: "Using API Carto"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{api_carto}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(happign)
library(tmap); tmap_mode("view")
```

# Presentation

APIs carto were developed to automatically retrieve certain spatial information required in administrative forms (e.g. for a building permit application, the identifier of the cadastral parcel can often be obtained directly from the address). The main advantage of these APIs is that they can be queried without spatial data. To do this with `get_wfs`, you would have to use an ECQL query with the `ecql_filter` argument which can be tricky.

`happign` implements APIs carto through `get_apicarto_*` functions

## API carto cadastre

Documentation : [https://apicarto.ign.fr/api/doc/cadastre](https://apicarto.ign.fr/api/doc/cadastre)

The API carto cadastre provides the following information :
 * the boundaries of a town (`type = "commune"`)
 * the parcel sections or divisions (`type = "section"` or `type = "division"`)
 * the cadastral parcels (`type = "cadastre"`)
 * information on non-vectorized parcels (`type = "localisant"`)

At least three parameters must be set :
 * `x` : An indication about the location. Could be a shape, an insee code or a departement code
 * `type` : What service do you want to use? (see above)
 * `source` : The data source `"PCI"` for "Parcellaire Express" or `"BDP"` for "BD Parcellaire". The BD Parcellaire product is a historical product that is no longer updated. It is therefore strongly recommended to use the Parcellaire Express product which is updated every six months.
 
All other parameters are used to refine the query.

### Usage

We’ll start with a simple example : retrieve borders of multiple town. 
Because `get_apicaro_cadastre` is a vectorized function, it's possible to set multiple insee code.
If you do not know insee codes, you can consult existing codes from the internal 
dataframe `cog_2022`.

```{r message=FALSE}
# all town starting with plou
plou_insee_code <- cog_2022[startsWith(cog_2022$LIBELLE, "Plou"), "COM"]
plou_borders <- get_apicarto_cadastre(plou_insee_code, type = "commune")

# result
tm_shape(plou_borders)+
   tm_borders(col = "black")
```

Another common case consists in recovering the geometry of the parcels from a "cadastral matrix extract". The latter lists for each owner all his built and unbuilt properties owned in a commune. It is a private information and to obtain one it is necessary to ask for an extract top the Center of the Land taxes. In this example a false simplified cadastral matrix is used.

```{r}
cad_mat <- data.frame(CODE_DEP = rep("29", 10),
                      CODE_COM = rep("158", 10),
                      SECTION = rep(c("AX", "AV"), each = 5),
                      N_PARC = c("0001","0002","0003","0004","0005",
                                 "0116","0117","0118","0119","0120"))
                               
parcels <- get_apicarto_cadastre(paste0(cad_mat$CODE_DEP, cad_mat$CODE_COM),
                                 section = cad_mat$SECTION,
                                 numero = cad_mat$N_PARC)


tm_shape(parcels)+
   tm_borders(col = "black")
                               
```



 

