---
title: "SCAN 25, SCAN 100 et SCAN OACI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SCAN 25, SCAN 100 et SCAN OACI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(happign)
```

## Brief presentation

Even if almost all IGN data are available to the general public, some layers are distributed within a stricter framework of use. The SCAN 25速, SCAN 100速 and SCAN OACI (and associated online services) are distributed under license conditions that can be consulted [here](https://geoservices.ign.fr/cgu-licences#licence      .
The final use in a professional or associative context is free without exception, according to the terms of this license.
The final use for the general public is free of charge up to a certain threshold, then paying, according to the appropriate general [conditions](https://geoservices.ign.fr/cgu-licences#cgu). All infos can be found [here](https://geoservices.ign.fr/services-web-issus-des-scans-ign).

## What are they ?

- The SCAN 25速 is a unique reference, recognized for its topographic richness, and offers an exemplary representation of the territory (1/25 000)

- The SCAN 100速 is a perfect compromise between a rather rich topographic representation and a global visualization of the territory (1/100 000

- The SCAN OACI corresponds to the aeronautical reference maps in the field of civil aviation

## How to use them with happign ?

These layers are WMS raster, so you just have to use the `get_wms_raster()` function with a personal apikey. To do this, you need to create an account on the geoservice site and then request a key in the "My keys" section of your account. It should look like this : "abc12efghi34j5k678lmnopq" (obvisouly, this is a fake key).

Then you choose the layer you want at the [Web services from IGN SCANs](https://geoservices.ign.fr/services-web-issus-des-scans-ign) and it's done !

```{r, eval = FALSE, echo = TRUE}
# Create or import a shape
shape <- st_polygon(list(matrix(c(-4.373937, 47.79859,
                                  -4.375615, 47.79738,
                                  -4.375147, 47.79683,
                                  -4.373937, 47.79859),
                                ncol = 2, byrow = TRUE)))
shape <- st_sfc(shape, crs = st_crs(4326))

my_key <- "abc12efghi34j5k678lmnopq"
scan25_name <- "SCAN25TOUR_PYR-JPEG_WLD_WM"

scan25 <- get_wms_raster(shape,
                         apikey = my_key,
                         layer_name = scan25_name,
                         resolution = NULL,   # To have full resolution
                         filename = "SCAN25") # This resource need to be downloaded on disk


```


