---
title: "API Carto"
author: "Paul Carteron"
date: "2025-09-02"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{API Carto}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
   collapse = TRUE,
   comment = "#>",
   fig.center = TRUE,
   fig.align = "center"
)
```

```{r setup}
library(happign)
library(dplyr)
library(sf)
library(tmap)
tmap_mode("view")
```

# Presentation

APIs carto were developed to automatically retrieve certain spatial information required in administrative forms. The main advantage of these APIs is that they can be queried without spatial data. To do this with `get_wfs`, you would have to use an ECQL query with the `ecql_filter`.

`happign` implements APIs carto through `get_apicarto_*` functions.

## API carto cadastre

Documentation : [https://apicarto.ign.fr/api/doc/cadastre](https://apicarto.ign.fr/api/doc/cadastre)

The API carto cadastre provides following informations :

* the boundaries of a town (`type = "commune"`)
* the parcel sections or divisions (`type = "section"`)
* the cadastral parcels (`type = "parcelle"`)
* information on non-vectorized parcels (`type = "localisant"`)

At least three parameters must be set :

* `x` : An indication about the location. Could be a shape, an insee code or a departement code
* `type` : What service do you want to use? (see above)
* `source` : The data source `"PCI"` for "Parcellaire Express" or `"BDP"` for "BD Parcellaire". The BD Parcellaire product is a historical product that is no longer updated (but quite even with better
geospatialisation). It is therefore strongly recommended to use the Parcellaire Express 
product which is updated every six months.
 
All other parameters are used to refine the query.

### Usage

We’ll start with a simple example : retrieve borders of a town.

```{r penmarch_borders, message=TRUE}

penmarch <- get_apicarto_cadastre("29158", type = "commune")

# result
tm_shape(penmarch)+
   tm_polygons(col = "black")
```

Because `get_apicaro_cadastre` is a vectorized function, it's possible to set multiple insee code.
If you do not know insee codes, you can consult existing codes from the internal dataframe `com_2025`. Below an exemple that find all commune starting with "ker"

```{r ker_borders, message=TRUE}
data("com_2025", package = "happign")

# all town starting with "KER", yes I'm coming from "La Bretagne"
ker_insee_code <- com_2025[startsWith(com_2025$NCC_COM, "KER"), "COM"]
ker_borders <- get_apicarto_cadastre(ker_insee_code, type = "commune")

# result
tm_shape(ker_borders)+
   tm_polygons(col = "black")
```

Another common case consists in recovering the geometry of the parcels from a "cadastral matrix extract". The latter lists for each owner all his built and unbuilt properties owned in a commune. It is a private information and to obtain one it is necessary to ask for an extract top the Center of the Land taxes. In this example a false simplified cadastral matrix is used.

```{r mat_cadastral, message=FALSE}
cad_mat <- data.frame(CODE_DEP = rep("29", 10),
                      CODE_COM = rep("158", 10),
                      SECTION = rep(c("AX", "AV"), each = 5),
                      N_PARC = c("0001","0002","0003","0004","0005",
                                 "0116","0117","0118","0119","0120"))

parcels <- get_apicarto_cadastre(paste0(cad_mat$CODE_DEP, cad_mat$CODE_COM),
                                 section = cad_mat$SECTION,
                                 numero = cad_mat$N_PARC,
                                 type = "parcelle")
tm_shape(parcels)+
   tm_borders(fill = "red")

```


```{r rpg, eval=FALSE, include=FALSE}
## API carto RPG

Documentation : [https://apicarto.ign.fr/api/doc/rpg](https://apicarto.ign.fr/api/doc/rpg)

The graphic parcel register (RPG) is a geographic information system (GIS) allowing the identification of agricultural parcels. It is useful to the authorities working on agricultural development and land management (community of municipalities, watershed syndicate, etc.) when they need information on the characterization and evolution of the agricultural area.

### Usage

Let's start by finding RPG from Penmarc'h town for years 2015. One of the limitations of the API carto RPG is that complex geometries, such as the precise contours of a commune, are not supported. To get around this problem, `get_apicarto_rpg` allows to simplify the shape with the `dTolerance` parameter. The method is illustrated below.

penmarch  <- get_apicarto_cadastre("29158", type = "commune")
rpg <- get_apicarto_rpg(penmarch, annee = 2015, dTolerance = 15)

# plot result
tm_shape(penmarch)+
   tm_borders()+
tm_shape(rpg)+
   tm_polygons("code_cultu")
```


```{r rpg_vect, eval=FALSE, include=FALSE}
The function is vectorized, which mean that query several years at the same time is possible. However, since 2014 the format of the RPG has changed. The function will therefore return a list with the parcels before and after 2014. Below an example showing evolution of RPG from 2013 to 2016.

all_years_rpg <- get_apicarto_rpg(penmarch, annee = 2013:2016, dTolerance = 15)

# plot evolution of rpg since 2010
tm_shape(penmarch)+
   tm_borders()+
tm_shape(all_years_rpg[["2013"]])+
   tm_polygons(group = "2013", "code_cultu")+
tm_shape(all_years_rpg[["2014"]])+
   tm_polygons(group = "2014", "code_cultu")+
tm_shape(all_years_rpg[["2015"]])+
   tm_polygons(group = "2015", "code_cultu")+
tm_shape(all_years_rpg[["2016"]])+
   tm_polygons(group = "2016", "code_cultu")
```


```{r rpg_pdt_car_poi, eval=FALSE, include=FALSE}

It is also possible to query from the culture code. Again, there are changes before and after 2014; [The user documentation](https://apicarto.ign.fr/api/doc/pdf/docUser_moduleRPG.pdf) explains:

"For RPG version 1 (before 2014) the value to be entered corresponds to the number contained in the CODE_CULTU attribute of the ILOTS_ANONYMS class with a value ranging from **01 to 28** (0 is mandatory for values less than 10). The list of crop codes can be consulted in [the content description of the RPG version 1.0](https://geoservices.ign.fr/sites/default/files/2021-07/DC_DL_RPG_1-0.pdf).
For the RPG version 2 the value to be entered corresponds to the trigram contained in the attribute CODE_CULTU of the class PARCELLES_GRAPHIQUES by respecting the case (upper case). The list of crop codes can be found in the [content description of the GPR version 2.0.](https://geoservices.ign.fr/sites/default/files/2021-07/DC_DL_RPG_2-0.pdf)"

For this example, let's find all potatoes, carrots and leeks.


rpg <- get_apicarto_rpg(penmarch,
                        annee = 2020,
                        code_cultu = c("PTC", "CAR", "POR"),
                        dTolerance = 15)

tm_shape(rpg)+
   tm_polygons("code_cultu",
               palette = c("orange", "springgreen", "gold"))
```

## API carto urbanism

Documentation : [https://apicarto.ign.fr/api/doc/gpu](https://apicarto.ign.fr/api/doc/gpu)

All layer available have the GPU are retrieved using `get_gpu_layers()`

```{r gpu_layer}
all <- names(get_gpu_layers())
document_urbanisme <- names(get_gpu_layers("du"))
```

### RNU

First of all, you can check if a commune is under the National Urbanism Regulation from is insee code. The RNU fully apply in communes that have neither a local map nor a local urban plan (PLU, PLUi) nor a document in replacement of a PLU.

```{r is_gpu}
is_rnu <- get_apicarto_gpu("29158", layer = "municipality")
is_rnu$is_rnu

# Penmarch isn't under the RNU and therefore has a document of urbanism

is_rnu <- get_apicarto_gpu("23004", layer = "municipality")
is_rnu$is_rnu

# Anzeme is under the RNU and therefore has a town planning document
```

### PLU, PLUi, POS, CC, PSMV

Urban planning documents can take several forms:
* PLU : Local Urbanism Plan
* PLUi : Intercommunal Local Urbanism Plan
* POS : Land use plan
* PSMV : Plan of Safeguarding and Development
* CC : Communal map

The first step is to find out if urban planning document are available if it is the case, find the document's partition i.e. its ID :

```{r get_urban_plan, message=FALSE}
# find out if documents are available
penmarch <- get_apicarto_cadastre("29158", "commune")
doc <- get_apicarto_gpu(st_centroid(penmarch), "document") 
# It's better to use centroid instead of borders to avoid conflict with other communes

partition <- doc$partition
```

Now that the partition is recovered, it is possible to obtain several resources for a specific document. The different resources available are specified in the documentation of the function `?get_apicarto_gpu()`

```{r zone_urba, message=FALSE}
zone_urba <- get_apicarto_gpu(partition, "zone-urba")

# click on polygon for legend
tm_shape(zone_urba)+
   tm_polygons("libelong", legend.show = FALSE)

```

Because `get_apicarto_gpu` is vectorized on `x`, many insee code or partition can be queried.

```{r gpu_vect_prescription, message=FALSE}
all_ker_gpu <- get_apicarto_gpu(ker_borders$code_insee, "municipality")
```

### SUP : Servitude d'utilité publique

GPU also give access to SUP databases which contain multiple category such as : protection forest, agricultural protected area, historical monument, ...)
See [national nomenclature](https://www.geoportail-urbanisme.gouv.fr/infos_sup/) for more info.

It give access to "assiette" and "generateur". First one is the restructed area, second one is what generated this area.

Partition of "assiette" and "generateur" are quite complicated to found so generally geometry is used.

```{r assiette_vs_generateur, message=FALSE}
penmarch <- get_apicarto_cadastre("29158")
assiette <- get_apicarto_gpu(penmarch, "assiette-sup-s")
generateur <- get_apicarto_gpu(penmarch, "generateur-sup-s")

tm_shape(penmarch)+
   tm_borders()+
tm_shape(assiette)+
   tm_fill("suptype")+
tm_shape(generateur)+
   tm_fill("firebrick")
```
When using SUP layers, you can filter by category. Here all historical monument of Penmarch.

```{r assiette_vs_generateur, message=FALSE}
mh <- get_apicarto_gpu(penmarch, "assiette-sup-s", "AC1")

tm_shape(penmarch)+
   tm_borders()+
tm_shape(mh)+
   tm_polygons()
```

All layer "assiette", "generateur", "info", "prescription" have three type : point, line, polygon. If you want to query them all you can loop over it.

```{r assiette_vs_generateur, message=FALSE}
all_generateur_layers <- names(get_gpu_layers("generateur"))
generateurs <- lapply(all_generateur_layers, \(x) get_apicarto_gpu(penmarch, x)) |>
   setNames(all_generateur_layers)

tm_shape(penmarch)+
   tm_borders()+
tm_shape(generateurs$`generateur-sup-p`)+
   tm_point()
tm_shape(generateurs$`generateur-sup-l`)+
   tm_lines()
tm_shape(generateurs$`generateur-sup-s`)+
   tm_polygons()
```
